---
title: "Data Visualization with ggplot2"
author: "Melise Edwards"
date: "9/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Now, we will switch to doing the ggplot2 data vis lesson using data carpentry website : https://datacarpentry.org/R-ecology-lesson/04-visualization-ggplot2.html


```{r}
library(tidyverse)
```

```{r message=FALSE, include=FALSE}
surveys_complete <- read_csv("data/surveys_complete.csv")
```

```{r}
ggplot(data = surveys_complete, mapping = aes(x=weight, y=hindfoot_length)) + geom_point()
```



#### the below method shows you how you can save the ggplot metrics(?) or what you want to graph, then use the "+" symbol to add whichever visualization or plot type you want to the data. Very cool! 
```{r}
surveys_plot <- ggplot(data = surveys_complete, 
                       mapping = aes(x = weight, y = hindfoot_length))

surveys_plot + geom_point()
```


## Challenge 1
#### install hexbin
#### then use geom_hex() function

```{r}
library(hexbin)
surveys_plot + 
  geom_hex()
```


#### use transparency (alpha) to avoid overplotting. Values of (alpha) range from 0 to 1 with lower values corresponding to more transparent colors. Can also use the aesthetic parameters (colour) and (fil).

links for more learning = https://ggplot2.tidyverse.org/reference/aes_colour_fill_alpha.html

```{r}
surveys_plot +
  geom_point(alpha=0.1)
```




#### adding colors 

```{r}
ggplot(surveys_complete, aes(x=weight, y=hindfoot_length)) +
  geom_point(alpha=0.1, color= "blue")
```


#### to add color by species id
```{r}
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length)) + 
  geom_point(alpha = 0.1, aes(color = species_id))
```


### Challenge 2
#### Use what you just learned to create a scatter plot of weight over species_id with the plot types showing in different colors. Is this a good way to show this type of data?

No this does not seem like the best way to show the data, maybe a split boxplot or something...

```{r}
ggplot(surveys_complete, aes(x = species_id, y = weight)) + 
  geom_point(aes(color = plot_type))
```



### Boxplots 

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
         geom_boxplot()
```


#### tomato colors?! this would be perfect for our cognitive testing scores and PCR differences in protein expression depending on treatment and/or sex. 

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
  geom_jitter(alpha = 0.3, color = "tomato") +
  geom_boxplot(alpha = 0)
 
```


### Challenge 3 
#### Replace the box plot with a violin plot; see geom_violin().

```{r}
violin.eg <- ggplot(data = surveys_complete, mapping = aes(x = species_id, y = weight)) + 
  geom_jitter(alpha = 0.3, color = "tomato") +
  geom_violin(alpha = 0) +
  theme_classic() + 
  ggtitle("Boxplot")

violin.eg + scale_y_log10()
```



#### try looking at something else, like hindfoot_length

```{r}
ggplot(data = surveys_complete, mapping = aes(x = species_id, y = hindfoot_length)) + 
  geom_jitter(alpha = 0.3, color = "cyan2") +
  geom_boxplot(alpha = 0) + 
  theme_classic()

demo("colors")
```




