---
title: "Lab 5 Extra"
author: "Melise"
date: "10/5/2020, late due to sickness"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Lab extra: Interactive Graphs and Animations from the COVID-19 reporting data

**Source:** https://moodle.umass.edu/pluginfile.php/3449324/mod_resource/content/4/EvoGeno_Lab5x_interactive_graphs.html

```{r}
#helpful tips for code and outputs in R markdown
# {r, code = FALSE} code will not be show, but will run and output will be shown 
# {r, eval = FALSE} code will be shown but will not run and no output 

library(tidyverse)
library(lubridate)
```

```{r}
time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                             names_to = "Date", values_to = "Confirmed") 
```

```{r}
# Let's get the times series data for deaths
time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")
```

```{r}
# Create Keys 
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)

```

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
head(time_series_deaths_long)
```

```{r}
# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
    time_series_deaths_long, by = c("Key")) %>% 
    select(-Key)
head(time_series_long_joined)
```

```{r}
# Reformat the data
# using the mdy function, it is now a date instead of a character and structured slightly different
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
head(time_series_long_joined)

```

```{r}
# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```


```{r}
# Plot graph to a pdf outputfile -- didn't work for me likely bc I do not have an images folder?

time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() 
```

```{r}
ppi <- 300
#unable to use png() function

time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Deaths")

```


```{r}
# Version 2
library(plotly)
```


```{r}

ggplotly(
  time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
      geom_point() +
      geom_line() +
      ggtitle("US COVID-19 Deaths")
 )

```


```{r}
# Subset the time series data to include US deaths
US_deaths <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region == "US")

head(US_deaths)
# Collect the layers for agraph of the US time series data for covid deaths
 p <- ggplot(data = US_deaths, aes(x = Date,  y = Deaths)) + 
        geom_point() +
        geom_line() +
        ggtitle("US COVID-19 Deaths")
# Plot the graph using ggplotly
ggplotly(p)
```



### Animated Graphs with gganimate
Animated graphs when down right have a great visual impact. You can do this in R and have your animations embedded on your web page. Essentially gganimate creates a series of files that are encompassed in a gif file. In addition to having this gif as part of your report file, you can save the gif and use in a slide or other presentations. It just takes a few lines of code to covert and existing ggplot graph into an animation. 

This are some important gganimate functions:

__transition_*()__ defines how the data should be spread out and how it relates to itself across time.
__view_*()__ defines how the positional scales should change along the animation.
__shadow_*()__ defines how data from other points in time should be presented in the given point in time.
enter_()/exit_() defines how new data should appear and how old data should disappear during the course of the animation.
ease_aes() defines how different aesthetics should be eased during transitions.


```{r, eval=FALSE}
# packages to install

install.packages("gganimate")
install.packages("transformr")
install.packages("gifski")
install.packages("av")

library(gganimate)
library(transformr)
theme_set(theme_classic())
```


```{r}
# An animation of the confirmed cases in select countries

data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","Italy","US")) 
head(data_time)

```


```{r, fig.width = 6, fig.height = 6}
library(gganimate)
p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
  transition_reveal(Date)
       
animate(p, end_pause = 15)
```



```{r}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Japan","Italy","US")) 


p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 

# Some people needed to use this line instead
# animate(p,renderer = gifski_renderer(), end_pause = 15)
anim_save("deaths_5_countries.gif", p)
# saved in genomics-course folder 
```



