---
title: "Lab11x"
author: "Melise Edwards"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 11x Micrombiome analysis using phyloseq 

```{r}
library(phyloseq); packageVersion("phyloseq")
```

```{r}
data(GlobalPatterns)
library("ggplot2"); packageVersion("ggplot2")
```
```{r}
library("plyr"); packageVersion("plyr")
```
```{r}
theme_set(theme_bw()) #set theme
```


## **Tutorial 1: Ordination**
* Note: pre-processing decisions are important and varied based on your data; important to document decisions and rationale as you code
* OPU = operational taxonomic unit 
* **Ordination**:Ordination is a collective term for multivariate techniques which adapt a multi-dimensional swarm of data points in such a way that when it is projected onto a two dimensional space any intrinsic pattern the data may possess becomes apparent upon visual inspection (Pielou, 1984).


```{r}
# remove OTUs that do not show appear more than 5 times in more than half the samples
GP = GlobalPatterns
wh0 = genefilter_sample(GP, filterfun_sample(function(x) x > 5), A=0.5*nsamples(GP))
GP1 = prune_taxa(wh0, GP)
```

```{r}
#transform to even sampling depth
GP1 = transform_sample_counts(GP1, function(x) 1E6 * x/sum(x))
```

```{r}
# keep only the most abundant 5 phyla 
phylum.sum = tapply(taxa_sums(GP1), tax_table(GP1)[, "Phylum"], sum, na.rm=TRUE)
top5phyla = names(sort(phylum.sum, TRUE))[1:5]
GP1 = prune_taxa((tax_table(GP1)[, "Phylum"] %in% top5phyla), GP1)
```

```{r}
# define a human-associated versus non-human categorical variable because we want to assess a major prior among samples -- aka some are human sample and some are not
human = get_variable(GP1, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
sample_data(GP1)$human <- factor(human)

```

******

### Four main Ordination plots 

#### 1. Just OTUs 

```{r}
# shade points by phylum
GP.ord <- ordinate(GP1, "NMDS", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Phylum", title="taxa")
print(p1)
```

```{r}
# the above looks complicated; too much data in one plot. Use facet wrap to isolate and get a better visual of the data 

p1 + facet_wrap(~Phylum, 5)

```

#### 2. Just Samples 

```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="SampleType", shape="human") 
p2 + geom_polygon(aes(fill=SampleType)) + geom_point(size=5) + ggtitle("samples")
```


### 3. biplot graphic

```{r}
#both samples and OTUs are plotted together in one "biplot"

p3 = plot_ordination(GP1, GP.ord, type="biplot", color="SampleType", shape="Phylum", title="biplot")

# Some stuff to modify the automatic shape scale
GP1.shape.names = get_taxa_unique(GP1, "Phylum")
GP1.shape <- 15:(15 + length(GP1.shape.names) - 1)
names(GP1.shape) <- GP1.shape.names
GP1.shape["samples"] <- 16
p3 + scale_shape_manual(values=GP1.shape)
```


### 4. split graphic 

```{r}
p4 = plot_ordination(GP1, GP.ord, type="split", color="Phylum", shape="human", label="SampleType", title="split") 
p4
```

```{r}
gg_color_hue <- function(n){
    hues = seq(15, 375, length=n+1)
    hcl(h=hues, l=65, c=100)[1:n]
}
color.names <- levels(p4$data$Phylum)
p4cols <- gg_color_hue(length(color.names))
names(p4cols) <- color.names
p4cols["samples"] <- "black"
p4 + scale_color_manual(values=p4cols)
```

****

### Supported ordination methods 

```{r}
dist = "bray"
ord_meths = c("DCA", "CCA", "RDA", "DPCoA", "NMDS", "MDS", "PCoA")
plist = llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples", color="SampleType")
}, GP1, dist)
```

```{r}
names(plist) <- ord_meths
```


```{r}
pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"
```


```{r}
p = ggplot(pdataframe, aes(Axis_1, Axis_2, color=SampleType, shape=human, fill=SampleType))
p = p + geom_point(size=4) + geom_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + scale_colour_brewer(type="qual", palette="Set1")
p
```

```{r}
plist[[2]] #prints the second element of the list 
```

```{r}
#now add some layers to make it look nicer 
p = plist[[2]] + scale_colour_brewer(type="qual", palette="Set1")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + geom_point(size=5) + geom_polygon(aes(fill=SampleType))
p
```


### MDS ("PCoA") on Unifrac distances 

```{r}
ordu = ordinate(GP1, "PCoA", "unifrac", weighted=TRUE)
plot_ordination(GP1, ordu, color="SampleType", shape="human")
```

```{r}
# more ggplot2 layers 

p = plot_ordination(GP1, ordu, color="SampleType", shape="human")
p = p + geom_point(size=7, alpha=0.75)
p = p + scale_colour_brewer(type="qual", palette="Set1")
p + ggtitle("MDS/PCoA on weighted-UniFrac distance, GlobalPatterns")
```

**************
*************

## Part II: Alpha Diversity 

```{r}
data("GlobalPatterns")
```


```{r}
theme_set(theme_bw())
pal = "Set1"
scale_colour_discrete <-  function(palname=pal, ...){
  scale_colour_brewer(palette=palname, ...)
}
scale_fill_discrete <-  function(palname=pal, ...){
  scale_fill_brewer(palette=palname, ...)
}

```


```{r}
GP <- prune_species(speciesSums(GlobalPatterns) > 0, GlobalPatterns)
```

```{r}
plot_richness(GP)
```

```{r}
plot_richness(GP, measures = c("Chao1", "Shannon"))
```

```{r}
plot_richness(GP, x="SampleType", measures = c("Chao1", "Shannon"))
```

```{r}
# the following few chunks are intending to introduce a logical to indicate whether or not the samples are human-associated

sampleData(GP)$human <- getVariable(GP, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
```

```{r}
plot_richness(GP, x="human", color="SampleType", measures=c("Chao1", "Shannon"))
```

```{r}
# can merge samples from environment 
GPst = merge_samples(GP, "SampleType")
# repair variables that were damaged during merge (coerced to numeric)
sample_data(GPst)$SampleType <- factor(sample_names(GPst))
sample_data(GPst)$human <- as.logical(sample_data(GPst)$human)
```


```{r}
p = plot_richness(GPst, x="human", color="SampleType", measures=c("Chao1", "Shannon"))
p + geom_point(size=5, alpha=0.2)
```

```{r}
# check which lists are present in p 
p$layers
```

```{r}
# we can use negative indexing to 'pop out' the first layer, which has small points? then add a new geom_point layer with larger point size

p$layers <- p$layers[-1]
p + geom_point(size = 5, alpha = 0.5)
```

**********
**********

## HEAT MAPS!! 
Following the [github tutorial](https://joey711.github.io/phyloseq/plot_heatmap-examples.html)

```{r}
data("GlobalPatterns")
gpt <- subset_taxa(GlobalPatterns, Kingdom=="Bacteria")
gpt <- prune_taxa(names(sort(taxa_sums(gpt), TRUE)[1:300]),gpt)
plot_heatmap(gpt, sample.label = "SampleType")
```

```{r}
#subset the data to something more manageable

gpac <- subset_taxa(GlobalPatterns, Phylum=="Crenarchaeota")
plot_heatmap(gpac)

```


```{r}
#relabel by 'SampleType' sample variable and the taxonomic rank of 'Family'
(p <- plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family"))
```

```{r}
# relab axes
p$scales$scales[[1]]$name <- "My X-Axis"
p$scales$scales[[2]]$name <- "My Y-Axis"
print(p)
```

```{r}
# change color scheme
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#000033", high="#CCFF66")
```

```{r}
# dark blue to red color scheme
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#000033", high="#FF3300")
```

```{r}
#dark blue to light blue 
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#000033", high="#66CCFF")
```

```{r}
# change background by changing "NA" and "0" elements 
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#66CCFF", high="#000033", na.value="white")
```

```{r}
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#FFFFCC", high="#000033", na.value="white")
```

*****

```{r}
# okay enough color schemes. Now try different ordination methods 

plot_heatmap(gpac, "NMDS", "jaccard")
```

```{r}
plot_heatmap(gpac, "DCA", "none", "SampleType", "Family")
```

```{r}
plot_heatmap(gpac, "RDA", "none", "SampleType", "Family") #PCA analysis 
```

```{r}
# PCoA/MDS ordination on the (default) bray-curtis distance
plot_heatmap(gpac, "PCoA", "bray", "SampleType", "Family")
```

```{r}
# MDS/PCoA ordination on the Unweighted-UniFrac distance.

plot_heatmap(gpac, "PCoA", "unifrac", "SampleType", "Family")
```

```{r}
# weighted-UniFrac dstance and MDS/PCoA ordination 
plot_heatmap(gpac, "MDS", "unifrac", "SampleType", "Family", weighted=TRUE)
```

```{r}
# Here is how you might create a heatmap using base-R graphics and the more common (but problematic) hierarchical clustering organization, in case you want to compare with plot_heatmap, for example.
heatmap(otu_table(gpac))
```

**********
*********

## Part IV: Networks 
Following the [phyloseq tutorial on github](http://joey711.github.io/phyloseq/plot_network-examples.html)


```{r}
data(enterotype)
```


```{r}
set.seed(711L)
```

```{r}
enterotype = subset_samples(enterotype, !is.na(Enterotype))
```


```{r}
plot_net(enterotype, maxdist = 0.4, point_label = "Sample_ID")
```


```{r}
plot_net(enterotype, maxdist = 0.3, color = "SeqTech", shape="Enterotype")
```

```{r}
#plot_network function 

ig <- make_network(enterotype, max.dist = 0.3)

plot_network(ig, enterotype)
```

```{r}
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

```{r}
# decrease number of edges in the network and lower distance from .3 to .2

ig <- make_network(enterotype, max.dist=0.2)
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

```{r}
ig <- make_network(enterotype, dist.fun="bray", max.dist=0.3)
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

