---
title: "Lab 3 - Data Wrangling"
author: "Melise Edwards"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Notes from Sept 7, 2020 Class 

### Calling upon tidyverse after installation 

```{r}
library(tidyverse)
library(crayon)
```

### downloading file from moodle

```{r}
download.file(url="https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")

```

### loading file

```{r}
surveys <- read.csv("data/portal_data_joined.csv")
```

surveys
```{r}
head(surveys)
```

### to open in another tab and see as a table, use the function view()
```{r}
view(surveys)
```


### inspecting the structure of a dataframe

```{r}
str(surveys)
```

### To get a sense of the structure of the dataframe we can use the following:
For size:
dim(surveys)
nrow(surveys)
ncol(surveys)

For content - first and last six rows:
head(surveys)
tail(surveys)

Names:
colnames(surveys)
rownames(surveys)

# Challenge 1

Based on the output of str(surveys), can you answer the following questions?

# *What is the class of the object surveys?*
```{r}
class(surveys)
```

# *How many rows and how many columns are in this object?*

```{r}
nrow(surveys)
```
```{r}
ncol(surveys)
```

# *How many species have been recorded during these surveys?*

```{r}
length(unique(surveys$species))
```

```{r}
length(unique(surveys$species_id))
```


### Inspecting the dataframe 

```{r}
surveys[3,]
```

```{r}
head_surveys <- surveys[1:6, ]
```

When you do surveys["species_id"]
This is a dataframe

When you do surveys[, "species_id"]
This however returns a vector, same thing as surveys$species_id 

# Challenge 2

## *1. Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.*

```{r}
surveys_200 <- surveys[200, ]
```

## *2. Notice how nrow() gave you the number of rows in a data.frame?*
Use that number to pull out just that last row in the data frame.
```{r}
nrow(surveys)
```

## *Compare that with what you see as the last row using tail() to make sure it’s meeting expectations.*

```{r}
tail(surveys)
```

## *Pull out that last row using nrow() instead of the row number.*

```{r}
n_row <- nrow(surveys)
```

## *Create a new data frame (surveys_last) from that last row.*

```{r}
surveys_last <- surveys[n_row, ]
```

## 3. *Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.*

```{r}
n_row/2
```

```{r}
surveys_middle <- surveys[17393, ]
```

```{r}
surveys_middle
```

## *4. Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.*

```{r}
surveys[-(7:n_row), ]
```

## New lesson on Factors 

```{r}
sex <- factor(c("male", "female", "female", "male"))
```

```{r}
levels(sex)
```

```{r}
nlevels(sex)
```

Examples

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
```

```{r}
as.numeric(levels(year_fct))
```

## Renaming factors lesson 

```{r}
plot(as.factor(surveys$sex))
```

```{r}
sex <- factor(surveys$sex)
```
```{r}
head(sex)
```

```{r}
levels(sex)
```

```{r}
levels(sex) [1] <- "undetermined"

```

```{r}
head(sex)
```

# Challenge 
Rename “F” and “M” to “female” and “male” respectively.
```{r}
levels(sex) [2] <- "Female"
```

```{r}
levels(sex) [3] <- "Male"
```

```{r}
levels(sex)
```


Now that we have renamed the factor level to “undetermined”, can you recreate the barplot such that “undetermined” is last (after “male”)?

```{r}
levels(sex)
```

```{r}
sex <- factor(sex, levels = c("Female", "Male", "undetermined"))
```

```{r}
plot(sex)
```

```{r}
surveys <- read.csv("data/portal_data_joined.csv", stringsAsFactors = TRUE)
surveys
```

```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
```

