---
title: "Lab3b - Manipulating the Data"
author: "Melise Edwards"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


*Fun fact:* the tidyverse package actually is an umbrella for several packages, which it installs all at once? 
Packages included = tidyr, dplyr, ggplot2, tibble, etc 

Link for today's exercise = https://datacarpentry.org/R-ecology-lesson/03-dplyr.html

```{r}
library("tidyverse")
```



#### Important note (likely important for RNAseq data) on the importance of dplyr:

An additional feature is the ability to work directly with data stored in an external database. The benefits of doing this are that the data can be managed natively in a relational database, queries can be conducted on that database, and only the results of the query are returned.

This addresses a common problem with R in that all operations are conducted in-memory and thus the amount of data you can work with is limited by available memory. The database connections essentially remove that limitation in that you can connect to a database of many hundreds of GB, conduct queries on it directly, and pull back into R only what you need for analysis.


tidyr - essential for data manipulation! 


```{r}
surveys <- read_csv("data/portal_data_joined.csv")
```

```{r}
## inspect the data
str(surveys)
```

##### can also preview the data using
View(surveys)


### using the select function in dplyr
*Note* --> This is going to be extremely helpful for qPCR data to take columns that are actually useful and then even save those columns as a new dataframe or tibble. 


```{r}
select(surveys, plot_id, species_id, weight)
```


To select all columns $except$ some of them, do this

```{r}
select(surveys, -record_id, -species_id)
```


```{r}
filter(surveys, year==1995)
```

Returns all the rows in which the year is 1995 (note: use filter, not select()) 


#### intermediate steps example as an alternative to piping

```{r}
surveys2 <- filter(surveys, weight < 5)
surveys_sml <- select(surveys2, species_id, sex, weight)
head(surveys_sml)

```



#### you can also nest functions like this 

```{r}
surveys_sml <- select(filter(surveys, weight < 5), species_id, sex, weight)
head(surveys_sml)
```



#### piping, %>% 


```{r}
surveys_sml <- surveys %>% 
  filter(weight < 5) %>% 
  select(species_id, sex, weight)

```


### Challenge: 
#### *Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight*

```{r}
challenge1 <- surveys %>% 
  filter(year < 1995) %>% 
  select(year, sex, weight)
head(challenge1)

```


#### Mutate 
Frequently you’ll want to create new columns based on the values in existing columns, for example to do unit conversions, or to find the ratio of values in two columns. For this we’ll use mutate().

```{r}
surveys %>%
  mutate(weight_kg = weight / 1000)
```


*This just created a new column of weight that is represented in kilograms*



```{r}
mutate.exp <- surveys %>%
  mutate(weight_kg = weight / 1000,
         weight_lb = weight_kg * 2.2)

head(mutate.exp)
```


*This just shows you that you can do two mutations in the same "call" 


#### instead of constantly saving new dataframes as an object in your environment to see the (head) try using pipes instead! 

```{r}
surveys %>%
  mutate(weight_kg = weight / 1000) %>%
  head()
```



```{r}
surveys %>% 
  filter(!is.na(weight)) %>% 
  mutate(weight_kg = weight / 1000) %>% 
  head()
```


### Challenge
#### *Create a new data frame from the surveys data that meets the following criteria: 
* contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. 
* In this hindfoot_cm column, there are no NAs and all values are less than 3.*


```{r}
surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  mutate(hindfoot_cm = hindfoot_length / 10) %>% 
  filter(hindfoot_cm < 3) %>% 
  select(hindfoot_cm, species_id) %>% 
  head()
  
```


#### Split-apply lesson


group_by() is often used together with summarize(), which collapses each group into a single-row summary of that group. group_by() takes as arguments the column names that contain the categorical variables for which you want to calculate the summary statistics. So to compute the mean weight by sex:

### this is so cool and necessary for the cognition data we have!!!!!!!!!!
```{r}
surveys %>%
  group_by(sex) %>%
  summarize(mean_weight = mean(weight, na.rm = TRUE))
```



```{r}
surveys %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight, na.rm = TRUE)) %>% 
  tail()
```



```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight)) %>%
  print(n = 15)
```



```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight),
            max_weight = max(weight))
```


#### arrange results in a specific order, neat! 


```{r}
 surveys %>% 
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(min_weight)
```


#### counting
```{r}
surveys %>% 
  count(sex)
```


#### another way to do this using group_by() and summarise ()


```{r}
surveys %>%
    group_by(sex) %>%
    summarise(count = n())
```



```{r}
surveys %>%
    count(sex, sort = TRUE) 
```


```{r}
surveys %>%
    count(sex, species)
```



### Challenge
#### *How many animals were caught in each plot_type surveyed?*

```{r}
surveys %>% 
  count(plot_type)
```


#### Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).


```{r}
surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  group_by(species_id) %>% 
  summarize(
    mean_hindfootlength = mean(hindfoot_length),
    max_hindfootlength = max(hindfoot_length),
    min_hindfootlength = min(hindfoot_length),
    n = n()
  ) %>% 
  head()
```


#### What was the heaviest animal measured in each year? Return the columns year, genus, species_id, and weight.

my answer first then theirs 

```{r}
surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(year) %>% 
  filter(weight == max(weight)) %>% 
  select(year, genus, species, weight) %>% 
  arrange(year) %>% 
  head()
```


```{r}
surveys %>%
    filter(!is.na(weight)) %>%
    group_by(year) %>%
    filter(weight == max(weight)) %>%
    select(year, genus, species, weight) %>%
    arrange(year) %>% 
  head()
```



#### spreading (needs extra practice for me)

```{r}
surveys_gw <- surveys %>%
  filter(!is.na(weight)) %>%
  group_by(plot_id, genus) %>%
  summarize(mean_weight = mean(weight))
```


```{r}
str(surveys_gw)
```


```{r}
surveys_gw %>%
  spread(genus, mean_weight, fill = 0) %>%
  head()
```


#### gathering

```{r}
surveys_gather <- surveys_gw %>%
  gather(key = "genus", value = "mean_weight", -plot_id)

```


### Challenge 
*Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function! See ?n_distinct for more.* 

```{r}
surveys_spread_genera <- surveys %>%
  group_by(plot_id, year) %>%
  summarize(n_genera = n_distinct(genus)) %>%
  spread(year, n_genera)
```

```{r}
head(surveys_spread_genera)
```



*Now take that data frame and gather() it again, so each row is a unique plot_id by year combination.* 


```{r}
surveys_spread_genera %>%
  gather("year", "n_genera", -plot_id)
```


*The surveys data set has two measurement columns: hindfoot_length and weight. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. Let’s walk through a common solution for this type of problem. First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight. Hint: You’ll need to specify which columns are being gathered.* 

```{r}
surveys_long <- surveys %>%
  gather("measurement", "value", hindfoot_length, weight)
```


*With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight. Hint: You only need to specify the key and value columns for spread().*

```{r}
surveys_long %>%
  group_by(year, measurement, plot_type) %>%
  summarize(mean_value = mean(value, na.rm=TRUE)) %>%
  spread(measurement, mean_value) %>% 
  head()
```


```{r}
surveys_complete <- surveys %>%
  filter(!is.na(weight),           # remove missing weight
         !is.na(hindfoot_length),  # remove missing hindfoot_length
         !is.na(sex))

species_counts <- surveys_complete %>%
    count(species_id) %>% 
    filter(n >= 50)

surveys_complete <- surveys_complete %>%
  filter(species_id %in% species_counts$species_id)


write_csv(surveys_complete, path = "data/surveys_complete.csv")
```

